### 2016-04-22 Manuscript/SW_early_life_telomeres.Rmd
DSR edtis

###Abstract
1. Understanding the costs of individual early-life experiences is fundamental to understanding life-history evolution. Telomeres, the protective caps at the ends of chromosomes, shorten in response to oxidative stress, and telomere shortening is correlated with reduced survival. Thus, it is possible that telomere dynamics may help us quantify individual variation in early-life costs, and enhance our understanding of how poor conditions in early life are related to later-life survival.

@@

3. We found that telomere length decreases with cross-sectional and longitudinal measures of age, and that telomere length decreases most rapidly early in life. However, at the individual level we found a large amount of variation in telomere length and rates of change, with no relationship between within-individual juvenile and adult telomere length.

4. Juvenile telomere length varied markedly among cohorts, but we found no evidence that this variation was related to temporal variation in food availability or population density. Within cohorts, we found some evidence that telomere dynamics were related to territory quality and the presence of helpers in the natal territory, but the majority of variation in juvenile telomere length was unexplained. We found no relationship between early-life telomere dynamics and survival to adulthood or later in life.

5. Our data suggest that although telomeres shorten with age, in natual popuations telomere dynamics vary enormously over both space and time. Ascertaining what explains this variation will be a difficult but worhtwhile task, requiring comprehensive, long-term ecological and quantitative genetic data.

@@

#####Study species and sampling
The Seychelles warbler is a small (~15 g), insectivorous passerine bird with a mean life expectancy of 5.5 years at fledging [@Hammers2013]. The population of *ca*. 320 birds on Cousin Island (04'20'S, 55'40'E) has been intensively studied since 1986 [@Richardson2003b; @Spurgin2014a]. This species has a main breeding season, running from June-August, when the breeding females on each of the *ca*. 115 territories lay one or, rarely, two or three eggs [@Komdeur1991]. As a result of this low reproductive output, combined with higher mortality in first-year birds [39%; @Brouwer2006], cohort sizes in the Seychelles warbler are small (<50).

@@

For a subset of first-year birds for which we had longitudinal data, with an additional sample taken as an adult. For these individuals we calculated the within-individual change in RTL by subtracting early-life RTL from adult RTL (hereafter $\Delta$RTL). Negative values of $\Delta$RTL reflect telomere shortening, while positive values reflect telomere lengthening.

#####Statistical analyses
We performed all analyses using R version 3.0.1 [@RDevelopmentCoreTeam2011]. Because we are using a different measure of telomere length to the previous study of adult telomere dynamics in this species [@Barrett2013a], as well as including many additional samples, we first re-tested, using all available samples, whether telomere length decreases with age and $\Delta$age [a longitudinal measure based on within-subject centring; @VandePol2009] in the Seychelles warbler, using linear mixed models with individual ID as a random effect. For the analysis of $\Delta$age we we only included birds who had been sampled at least twice. We then explored how telomere length varied within the first year of life (telomere length vs age in months) and among cohorts (telomere length vs cohort ID), again using linear mixed models with individual ID as a random effect. We tested how temporal variation in telomere length varied with temporal variation in population density (measured as the number of adult birds on the island at the end of each breeding season) and food availability, using linear mixed models with cohort ID as a random effect.

@@

We used generalised linear mixed models to test whether telomere length/loss were related to survival to adulthood. Models were set up with a binomial error structure, and included RTL/$\Delta$RTL and juvenile age as explanatory variable, along with the same random effects as for the model averaging (see above). We then used Cox regression, implemented in the survival package in R [@Therneau2011], to test whether long-term survival was related to individual-level telomere dynamics. For the cross-sectional data we used lifespan (in years) as survival time, while for the longitudinal data we used remaining lifespan from the second sampling event. We ran these models with cohort as a frailty term to exclude cohort effects, and only included cohorts in which at least 70 percent of individuals had died (all years up to and including 2007).

@@

#####Early-life RTL
Considering the entire Seychelles warbler lifespan, RTL decreased with age (LMM: t = `r round2(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID))$tTable[2,4])`; P `r round3(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID))$tTable[2,5])`), although there were clear increases in cross-sectional telomere length after one year of age and at several points later in life (Fig. 1A). RTL also decreased with age within individuals (i.e. with $\Delta$age; t = `r round2(summary(lme(RTL~DeltaAge,data=subset(dd,DeltaAge!=0),random=~1|BirdID))$tTable[2,4])`; P `r round3(summary(lme(RTL~DeltaAge,data=subset(dd,DeltaAge!=0),random=~1|BirdID))$tTable[2,5])`; Fig. 1B), and with age within the first year of life (i.e. excluding adult samples; t = `r round2(summary(lme(RTL~Agemonths,data=juv,random=~1|BirdID))$tTable[2,4])`; P `r round3(summary(lme(RTL~Agemonths,data=juv,random=~1|BirdID))$tTable[2,5])`).

Early-life RTL varied significantly among cohorts (Likelihood ratio test: F = `r round2(anova(lme(RTL~factor(FieldPeriodID),data=juv,random=~1|BirdID))[2,3])`; P `r round3(anova(lme(RTL~Agemonths,data=juv,random=~1|BirdID))[2,4])`). Cohort-level variation in juvenile RTL was positively related to adult population density (LMM controlling for age and cohort: t = `r round2(summary(lme(RTL~Agemonths+Density,data=juv,random=~1|FieldPeriodID))$tTable[3,4])`; P `r round3(summary(lme(RTL~Agemonths+Density,data=juv,random=~1|FieldPeriodID))$tTable[3,5])`; Figs 2B,2C). Variation in cohort-level RTL was not related temporal variation in  food availability (t = `r round2(summary(lme(RTL~Agemonths+Insect,data=juv,random=~1|FieldPeriodID,na.action=na.exclude))$tTable[3,4])`; P `r round3(summary(lme(RTL~Agemonths+Insect,data=juv,random=~1|FieldPeriodID,na.action=na.exclude))$tTable[3,5])`).

Within cohorts, the top model explaining variation in early-life RTL contained age, sex, tarsus length, and the interaction between age and tarsus length (Table S1). The top model was much better supported than a null model ($\Delta$AICc = `r round2(AICc(lmer(RTL~1+(1|FieldPeriodID)+(1|TerritoryID),data=juv,REML = FALSE))-AICc(lmer(RTL~Agemonths+Sex+Tarsus+Agemonths*Tarsus+(1|FieldPeriodID)+(1|TerritoryID),data=juv,REML = FALSE)))`), and with the exception of the sex and tarsus effects, model-averaged confidence intervals did not overlap zero (Fig. 3A). Tarsus length was negatively related to RTL in nestlings, unrelated to RTL in fledglings, and positively related to RTL in subadults (Fig. 3B). Males had slightly longer telomeres than females, particularly at the fledgling and subadult stages, although this effect was weak  (Fig. 3C).

#####Early-life telomere shortening
Longitudinal data showed that early-life RTL was not related to adult RTL (R^2^ = `r round3(summary(lm(RTL1~RTL+Agemonths,data=Loss))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=Loss))$coefficients[2,3])`; P = `r round3(summary(lm(RTL1~RTL+Agemonths,data = Loss))$coefficients[2,4])`; Fig. 4A). $\Delta$RTL was significantly lower than zero in nestlings (t = `r round2(t.test(subset(Loss,Agemonths == 1)$Loss)$statistic)`; P `r round3(t.test(subset(Loss,Agemonths == 1)$Loss)$p.value)`), but not in fledglings (t = `r round2(t.test(subset(Loss,Agemonths == 6)$Loss)$statistic)`; P = `r round2(t.test(subset(Loss,Agemonths == 6)$Loss)$p.value)`) or subadults (t = `r round2(t.test(subset(Loss,Agemonths == 10)$Loss)$statistic)`; P = `r round2(t.test(subset(Loss,Agemonths == 10)$Loss)$p.value)`). In line with this, $\Delta$RTL increased with age in early life (R^2^ = `r round2(summary(lm(Loss~Agemonths,data=Loss))$r.squared)`; t = `r round2(summary(lm(Loss~Agemonths,data=Loss))$coefficients[2,3])`; P `r round3(summary(lm(Loss~Agemonths,data = Loss))$coefficients[2,4])`; Fig. 4B), suggesting that telomere shortening was highest shortly after the nestling phase.

$\Delta$RTL did not differ significantly among cohorts (F = `r round2(anova(lm(Loss~factor(FieldPeriodID),data = Loss))[1,4])`; P `r round3(anova(lm(Loss~factor(FieldPeriodID),data = Loss))[1,5])`), although power for this analysis was limited as longitudinal sample sizes within cohorts were low. Model selection revealed that the top model explaining $\Delta$RTL within cohorts contained only age (Table S2). This model was a better fit than the null model ($\Delta$AICc = `r round2(AICc(lmer(Loss~1+(1|FieldPeriodID)+(1|TerritoryID),data=Loss,REML = FALSE))-AICc(lmer(Loss~Agemonths+(1|FieldPeriodID)+(1|TerritoryID),data=Loss,REML = FALSE)))`), and only the age effect was significant (Fig. 4C).

@@

Early-life RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~RTL+(1|FieldPeriodID),data=subset(juv13,Agemonths ==1),family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,4])`), with no interaction effect between RTL and juvenile age (estimate = `r round2(summary(glmer(SurvivedNext~RTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glmer(SurvivedNext~RTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[4,4])`). To test whether cohort-level effects on RTL influenced survival we calculated cohort-level RTL (i.e. the median RTL for each cohort) and added this term to a GLMM, and found that cohort-level RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~cohortTL+RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~cohortTL+RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,4])`). Using the longitudinal data, we found that $\Delta$RTL had no effect on survival to the year after the second sampling event (estimate = `r round2(summary(glm(SurvivedNext2~Loss+Agemonths,data = subset(Loss,LayYear<2013),family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glm(SurvivedNext2~Loss+Agemonths,data=subset(Loss,LayYear<2013),family = 'binomial'))$coefficients[2,4])`).

Cox regression showed no effect of individual-level or cohort-level RTL on survival later in life (individual-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~RTL+Agemonths+frailty(FieldPeriodID))))$coefficients[1,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~RTL+Agemonths+frailty(FieldPeriodID))))$coefficients[1,6])`; cohort-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL*Agemonths+RTL+frailty(FieldPeriodID))))$coefficients[1,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL*Agemonths+RTL+frailty(FieldPeriodID))))$coefficients[1,6])`), and no interaction between RTL and age on survival (individual-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~RTL*Agemonths+frailty(FieldPeriodID))))$coefficients[4,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~RTL*Agemonths+frailty(FieldPeriodID))))$coefficients[4,6])`). $\Delta$RTL was not related to longer-term survival (estimate `r round3(summary(coxph(with(subset(Loss,LayYear<2008),Surv(RemainingLife2+1,Died)~Loss+Agemonths)))$coefficients[1,1])`, P = `r round2(summary(coxph(with(subset(Loss,LayYear<2008),Surv(RemainingLife2+1,Died)~Loss)))$coefficients[1,5])`).

###Discussion
Here we use one of the largest datasets to date to assess the relationships between early-life conditions, telomere length and survival in a closed population of Seychelles warblers. We found that, as in other studies, telomeres are lost at the greatest rate  early in life. We found that telomere dynamics are subject to strong cohort effects. Within cohorts, we found no evidence that the ecological or social environment influenced telomere dynamics, and we found no evidence that early-life telomere length is related to juvenile or later-life survival.

@@

Despite our clear finding that RTL decreases with age throughout the Seychelles warbler lifespan (Fig. 1), we found that at the individual-level, juvenile and adult telomere were not correlated, and that telomere shortening can only be detected very early in life (Fig. 4). This is in contrast to other avian studies where birds were reared in laboratory [@Heidinger2012] or nest-box [@Boonekamp2014] conditions, in which juvenile and adult telomere length were highly correlated, and individual-level telomere shortening clearly occurs throughout the juvenile period, and into adulthood. However, in accordance with our study, a recent study on Soay sheep (*Ovis aries*) found only a very weak correlation between juvenile and adult RTL, and a highly complex pattern of telomere dynamics across the lifespan [@Fairlie2016]. It is too soon to speculate as to what may be generating this variation among species, but one possibility is that environmental variability itself affects the resolution with which we can observe telomere shortening in natural populations [see also @Fairlie2016]. Regardless, our data suggest the efficacy of telomeres as individual biomarkers of cost in wild populations will depend on the age structure of the study population, as well as the sampling resolution.

>>>>

###Abstract
1. Understanding the costs of individual early-life experiences is fundamental to understanding life-history evolution and population dynmamics. Telomeres, the protective caps at the ends of chromosomes, shorten in response to oxidative stress, and telomere shortening is correlated with reduced survival. Thus, telomere dynamics may help us quantify individual variation in early-life costs, and enhance our understanding of how poor conditions in early life are related to later-life survival.

@@

3. We found that telomere length decreases with cross-sectional and longitudinal measures of age, and that telomere length decreases most rapidly early in life. Juvenile telomere length varied markedly among cohorts, but we found no evidence that this variation was related to temporal variation in food availability or population density.

4. At the individual level we found a large amount of variation in telomere length and rates of change, with no relationship between within-individual juvenile and adult telomere length. We found weak evidence that telomere dnegatively related to telomere length, but the majority of variation in juvenile telomere length and dynamics was unexplained. We found no relationship between early-life telomere dynamics and survival to adulthood or later in life.

5. Our comprehensive, long-term data suggest that although telomeres shorten with age, in natual popuations telomere dynamics vary enormously over both space and time. Ascertaining what explains this variation will be a difficult but worhtwhile task, which will require combining long-term ecological and quantitative genetic data.

@@

#####Study species and sampling
The Seychelles warbler is a small (~15 g), insectivorous passerine bird with a mean life expectancy of 5.5 years at fledging [@Hammers2013]. The population of *ca*. 320 birds on Cousin Island (04'20'S, 55'40'E) has been intensively studied since 1986 [@Richardson2003b; @Spurgin2014a]. This species' main breeding season runs from June-August, when the breeding females on each of the *ca*. 115 territories lay one or, rarely, two or three eggs [@Komdeur1991]. As a result of this low reproductive output, combined with higher mortality in first-year birds [39%; @Brouwer2006], cohort sizes in the Seychelles warbler are small (<50).

@@

For a subset of first-year birds we had longitudinal data, with an additional sample taken as an adult. For these individuals we calculated the within-individual change in RTL by subtracting early-life RTL from adult RTL (hereafter $\Delta$RTL). Negative values of $\Delta$RTL reflect telomere shortening, while positive values reflect telomere lengthening.

#####Statistical analyses
We performed all analyses using R version 3.0.1 [@RDevelopmentCoreTeam2011]. Our dataset differs from slightly from that of Barrett *et al.* [-@Barrett2013a] in that we are using RTL calculated driectly from the qPCR data rather than converting into an absolute measure , and we are including many additional samples. We therefore first re-tested, using all available samples, whether telomere length decreases with age and $\Delta$age [a longitudinal measure based on within-subject centring; @VandePol2009] in the Seychelles warbler, using linear mixed models with individual ID as a random effect. For the analysis of $\Delta$age we we only included birds who had been sampled at least twice. We then, using the juvenile samples, explored how telomere length varied within the first year of life (telomere length vs age in months) and among cohorts (telomere length vs cohort ID), again using linear mixed models with individual ID as a random effect. We tested how temporal variation in telomere length varied with temporal variation in population density (measured as the number of adult birds on the island at the end of each breeding season) and food availability, using linear mixed models with cohort ID as a random effect.

@@

We used generalised linear mixed models to test whether telomere length/loss were related to survival to adulthood. Models were set up with a binomial error structure, and included RTL or $\Delta$RTL and juvenile age as explanatory variables, along with the same random effects as for the model averaging (see above). We then used Cox regression, implemented in the survival package in R [@Therneau2011], to test whether long-term survival was related to individual-level telomere dynamics. For the cross-sectional data we used lifespan (in years) as survival time, while for the longitudinal data we used remaining lifespan from the second sampling event. We ran these models with cohort as a frailty term to exclude cohort effects, and only included cohorts in which at least 70 percent of individuals had died (all years up to and including 2007).

@@

#####Early-life RTL
Using the cross-sectional data RTL decreased with age across the entire lifespan of the Seychelles warbler (LMM: t = `r round2(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID))$tTable[2,4])`; P `r round3(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID))$tTable[2,5])`), although there were clear increases in (cross-sectional) telomere length after one year of age and at several points later in life (Fig. 1A). RTL also decreased with age within the first year of life (i.e. excluding adult samples; t = `r round2(summary(lme(RTL~Agemonths,data=juv,random=~1|BirdID))$tTable[2,4])`; P `r round3(summary(lme(RTL~Agemonths,data=juv,random=~1|BirdID))$tTable[2,5])`). In the longitudinal analysis RTL decreased with age within individuals (i.e. with $\Delta$age; t = `r round2(summary(lme(RTL~DeltaAge,data=subset(dd,DeltaAge!=0),random=~1|BirdID))$tTable[2,4])`; P = `r round3(summary(lme(RTL~DeltaAge,data=subset(dd,DeltaAge!=0),random=~1|BirdID))$tTable[2,5])`; Fig. 1B).

Early-life RTL varied significantly among cohorts (Likelihood ratio test: F = `r round2(anova(lme(RTL~factor(FieldPeriodID),data=juv,random=~1|BirdID))[2,3])`; P `r round3(anova(lme(RTL~Agemonths,data=juv,random=~1|BirdID))[2,4])`). Cohort-level variation in juvenile RTL was not related to adult population density (LMM controlling for age and cohort: t = `r round2(summary(lme(RTL~Agemonths+Density,data=juv,random=~1|FieldPeriodID))$tTable[3,4])`; P = `r round3(summary(lme(RTL~Agemonths+Density,data=juv,random=~1|FieldPeriodID))$tTable[3,5])`), or to temporal variation in  food availability (t = `r round2(summary(lme(RTL~Agemonths+Insect,data=juv,random=~1|FieldPeriodID,na.action=na.exclude))$tTable[3,4])`; P = `r round3(summary(lme(RTL~Agemonths+Insect,data=juv,random=~1|FieldPeriodID,na.action=na.exclude))$tTable[3,5])`).

Within cohorts, the top model explaining variation in early-life RTL contained age, territory quality, tarsus length, and the interaction between age and tarsus length (Table S1). The top model was much better supported than a null model ($\Delta$AICc = `r round2(AICc(lmer(RTL~1+(1|FieldPeriodID)+(1|TerritoryID),data=juv,REML = FALSE))-AICc(lmer(RTL~Agemonths+TQ+Tarsus+Agemonths*Tarsus+(1|FieldPeriodID)+(1|TerritoryID),data=juv,REML = FALSE)))`). Model-averaged confidence intervals of the age and tarsus x age effects did not overlap zero, but the tarsus and territory qusality effects were not significant (Fig. 3A). Tarsus length was negatively related to RTL in nestlings, unrelated to RTL in fledglings, and positively related to RTL in subadults (Fig. 3B). Territory quality was negatively related to RTL, particularly in nestlings, although this effect was very weak (Fig. 3C).

#####Early-life telomere shortening
Longitudinal data showed that early-life RTL was not related to adult RTL (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=Loss))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=Loss))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = Loss))$coefficients[2,4])`; Fig. 4A). Running this analysis separetly for juvenile age classes revealed that this relationship was not significant for nestlings (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 1)))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 1)))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = subset(Loss,Agemonths == 1)))$coefficients[2,4])`), fledglings (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 6)))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 6)))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = subset(Loss,Agemonths == 6)))$coefficients[2,4])`) or subadults (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 10)))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 10)))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = subset(Loss,Agemonths == 10)))$coefficients[2,4])`). $\Delta$RTL was significantly lower than zero in nestlings (t = `r round2(t.test(subset(Loss,Agemonths == 1)$Loss)$statistic)`; P `r round3(t.test(subset(Loss,Agemonths == 1)$Loss)$p.value)`), but not in fledglings (t = `r round2(t.test(subset(Loss,Agemonths == 6)$Loss)$statistic)`; P = `r round2(t.test(subset(Loss,Agemonths == 6)$Loss)$p.value)`) or subadults (t = `r round2(t.test(subset(Loss,Agemonths == 10)$Loss)$statistic)`; P = `r round2(t.test(subset(Loss,Agemonths == 10)$Loss)$p.value)`). In line with this, $\Delta$RTL increased with age in early life (R^2^ = `r round2(summary(lm(Loss~Agemonths,data=Loss))$r.squared)`; t = `r round2(summary(lm(Loss~Agemonths,data=Loss))$coefficients[2,3])`; P `r round3(summary(lm(Loss~Agemonths,data = Loss))$coefficients[2,4])`; Fig. 4B), suggesting that telomere shortening was highest shortly after the nestling phase.

$\Delta$RTL did not differ significantly among cohorts (F = `r round2(anova(lm(Loss~factor(FieldPeriodID),data = Loss))[1,4])`; P `r round3(anova(lm(Loss~factor(FieldPeriodID),data = Loss))[1,5])`), although power for this analysis was limited as longitudinal sample sizes within cohorts were low. Model selection revealed that the top model explaining $\Delta$RTL within cohorts contained only age (Table S2). This model was a better fit than the null model ($\Delta$AICc = `r round2(AICc(lmer(Loss~1+(1|FieldPeriodID)+(1|TerritoryID),data=Loss,REML = FALSE))-AICc(lmer(Loss~Agemonths+(1|FieldPeriodID)+(1|TerritoryID),data=Loss,REML = FALSE)))`), and only the age effect was significant over the top model set (Fig. 4C).

@@

Early-life RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~cenTL+Agemonths+(1|PlateID)+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~cenTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,4])`), with no interaction effect between RTL and juvenile age (estimate = `r round2(summary(glmer(SurvivedNext~cenTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glmer(SurvivedNext~cenTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[4,4])`), suggesting that there were no age-specific effects of juvenile RTL on survival to adulthood. To test whether cohort-level effects on RTL influenced survival we calculated cohort-level RTL (i.e. the median RTL for each cohort) and added this term to a GLMM, and found that cohort-level RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~cohortTL+RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~cohortTL+RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,4])`). Using the longitudinal data, we found that $\Delta$RTL had no effect on survival to the year after the second sampling event (estimate = `r round2(summary(glm(SurvivedNext2~Loss+Agemonths,data = subset(Loss,LayYear<2013),family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glm(SurvivedNext2~Loss+Agemonths,data=subset(Loss,LayYear<2013),family = 'binomial'))$coefficients[2,4])`).

Cox regression showed no effect of individual-level or cohort-level RTL on survival later in life (individual-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cenTL+Agemonths+frailty(FieldPeriodID))))$coefficients[1,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cenTL+Agemonths+frailty(FieldPeriodID))))$coefficients[1,6])`; cohort-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL*Agemonths+RTL+frailty(FieldPeriodID))))$coefficients[1,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL*Agemonths+RTL+frailty(FieldPeriodID))))$coefficients[1,6])`), and no interaction between RTL and juvenile age on survival (individual-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cenTL*Agemonths+frailty(FieldPeriodID))))$coefficients[4,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cenTL*Agemonths+frailty(FieldPeriodID))))$coefficients[4,6])`). $\Delta$RTL was not related to longer-term survival (estimate `r round3(summary(coxph(with(subset(Loss,LayYear<2008),Surv(RemainingLife2+1,Died)~Loss+Agemonths)))$coefficients[1,1])`, P = `r round2(summary(coxph(with(subset(Loss,LayYear<2008),Surv(RemainingLife2+1,Died)~Loss)))$coefficients[1,5])`).

###Discussion
Here we use one of the largest datasets to date to assess the relationships between early-life conditions, telomere length and survival in a closed population of Seychelles warblers. We found that, as in other studies, telomeres are lost at the greatest rate early in life. We found that telomere dynamics are subject to strong cohort effects. Within cohorts, we found little evidence that the ecological or social environment influenced telomere dynamics, and we found no evidence that early-life telomere length is related to juvenile or later-life survival.

@@

 We found that at the individual-level, juvenile and adult telomere were not correlated, and that consistent telomere shortening only occurs  very early in life (Fig. 4). This is in contrast to other avian studies where birds were reared in laboratory [@Heidinger2012] or nest-box [@Boonekamp2014] conditions, in which juvenile and adult telomere length were highly correlated, and individual-level telomere shortening clearly occurs throughout the juvenile period, and into adulthood. However, in accordance with our study, a recent study on Soay sheep (*Ovis aries*) found only a very weak correlation between juvenile and adult RTL, and a highly complex pattern of telomere dynamics across the lifespan [@Fairlie2016]. It is too soon to speculate as to what may be generating this variation among species, but one possibility is that environmental variability itself affects the resolution with which we can observe telomere shortening in natural populations [see also @Fairlie2016]. Regardless, our data suggest the efficacy of telomeres as individual biomarkers of cost in wild populations will depend on the age structure of the study population, as well as the sampling resolution.

### 2016-04-22 Manuscript/SW_early_life_telomeres_SOM.Rmd
DSR edtis

#Run the model
lm1 <- lmer(Loss ~ Agemonths*(Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|FieldPeriodID),
            data = temp,
            REML = FALSE)
lm1 <- standardize(lm1)

>>>>

#Run the model
lm1 <- lmer(Loss ~ Agemonths*(Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|FieldPeriodID) + (1|TerritoryID),
            data = temp,
            REML = FALSE)
lm1 <- standardize(lm1)
